---
title: "Vote Shiny"
author: "David Griswold"
date: "7/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r}
library(tidyverse)
library(jsonlite)
library(rvest)
library(magrittr)
```

# Cencus Data

## Read Data
```{r}
rawdemographicsData <- read_excel("poupByAge.xls")
```

## Create New Name Vector
```{r}
stateNames <- rawdemographicsData$STATE[!is.na(rawdemographicsData$STATE)]

correctStateNames <- c()

for (i in stateNames){
  correctStateNames <- c(correctStateNames, rep(i, 6))
}

rawdemographicsData$STATE <- correctStateNames

demographicsData <- rawdemographicsData

rm(i, correctStateNames, stateNames)
```

## Filter Our Data
```{r}
demographicsData %<>% filter(Age == "18 to 24")
```

## Delete Unecessary Columns
```{r}
demographicsData %<>% select(c("STATE", "Total Citizen Population", "Total registered", "Total voted")) 

names(demographicsData) <- c("State", "Poupulation", "Regestered_Voters", "Total_Voted")
```

## Type Conversions
```{r}
# Types good to go, will lowercase the state names
demographicsData$State %<>% str_to_title()
```

# Varible Manipulation
```{r}
demographicsData %<>% mutate(Percent_Regest = round(Regestered_Voters / Poupulation * 100, digits =2), 
                            Percent_Voting_Poup = round(Total_Voted / Poupulation * 100, digits =2),
                            Percent_Voting_Regest = round(Total_Voted / Regestered_Voters* 100, digits =2),
                            Num_Not_Voting = round(Poupulation * (1 - Percent_Voting_Poup / 100)))

```

## Drop US
```{r}
nationalData <- demographicsData[demographicsData$State == "Us",]


demographicsData <- demographicsData[!demographicsData$State == "Us",]
```

# Write Out
```{r}
write_csv(demographicsData, "demographicsData.csv")
```


# Scraping Voting Results

## Pull in Webpage Data
```{r}
base_url <- "https://www.politico.com/mapdata-2016/2016-election/results/map/president/"

ourElements <- ".timeline-group"

webpage <- read_html(base_url) %>% html_nodes(ourElements)
```

## Create Function for Cleaning Data
```{r}
cleanRow <- function (node) {
  
  stateName <- node %>% html_nodes(".timeline-header") %>% html_text() %>% str_trim()
  
  voteData <- node %>% html_node(".results-table") %>% html_table(fill = TRUE)
  
  return(list(state = stateName, data = voteData))
  
}
```


## Loop Function Over Function
```{r}
cleanedData <- map(webpage, function (x) cleanRow(x))
  
State <- map_chr(cleanedData, function(x) x$state)

clintonVotes <- map_chr(cleanedData, function(x) {
  
  ourData <- x$data
  
  index <- str_which(ourData$X1, "Clinton")
  
  return(ourData$X3[index])
})

trumpVotes <- map_chr(cleanedData, function(x) {
  
  ourData <- x$data
  
  index <- str_which(ourData$X1, "Trump")
  
  return(ourData$X3[index])
})


clintonPercent <- map_chr(cleanedData, function(x) {
  
  ourData <- x$data
  
  index <- str_which(ourData$X1, "Clinton")
  
  return(ourData$X2[index])
})

trumpPercent <- map_chr(cleanedData, function(x) {
  
  ourData <- x$data
  
  index <- str_which(ourData$X1, "Trump")
  
  return(ourData$X2[index])
})

voteData <- data.frame(State, 
                       clintonVotes,
                       trumpVotes,
                       clintonPercent,
                       trumpPercent)


```

## Clean Frame
```{r}
voteData$clintonVotes %<>% str_replace_all(",", "") %>% as.numeric()
voteData$trumpVotes %<>% str_replace_all(",", "") %>% as.numeric()

voteData$clintonPercent %<>% str_replace_all("%", "") %>% as.numeric()
voteData$trumpPercent %<>% str_replace_all("%", "") %>% as.numeric()

voteData$State %<>% as.character() %>% str_to_title()
```

## Add In Row for 

## Write Out To CSV
```{r}
write_csv(voteData, "PresVotesByState2016.csv")
```

# Combind Frames

## Merge Sets
```{r}
overallDataset <- inner_join(voteData, demographicsData, by = "State") %>%
   rbind(c("Puerto Rico",rep(0, 10))) # Adding a row for Puerto Rico
```

## Change Types
```{r}
states <- select(overallDataset, State)
overallDataset <- bind_cols(states, map_df(overallDataset[,2:ncol(overallDataset)], as.numeric))
```

## Add Varibles
```{r}
# Add Margin
overallDataset <- overallDataset %>% 
  mutate(Margin = round(trumpPercent - clintonPercent,2),
         GenZ_Power = abs(Margin) * Percent_Voting_Poup)
```

## Write Out  
```{r}
write_csv(overallDataset, file.path("Shiny_Vote", "VoteAndYouthTurnoutData2016.csv"))
```


